```{r}
NEI <- readRDS("./data/summarySCC_PM25.rds")
SCC <- readRDS("./data/Source_Classification_Code.rds")
library(ggplot2)
```

1. Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Make a plot showing the **total** PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r}
total_emissions_per_year <- aggregate(NEI$Emissions, list(NEI$year), sum)
colnames(total_emissions_per_year) <- c("Year", "Total")
total_emissions_per_year
ggplot(total_emissions_per_year, aes(x=Year,y=Total)) + labs(x="Year", y="Total PM2.5 emission from all sources", title="Total PM2.5 emission from all sources for each year") + geom_bar(width=0.8, stat="identity", fill="#5b23ad")
```

**Total emissions decreased from 1999 to 2008**

2. Have total emissions from PM2.5 decreased in the **Baltimore City**, Maryland (`fips == "24510"`) from 1999 to 2008?

```{r}
total_emissions_per_year_balt <- aggregate(NEI$Emissions[NEI$fips=="24510"], list(NEI$year[NEI$fips=="24510"]), sum)
colnames(total_emissions_per_year_balt) <- c("Year", "Total")
total_emissions_per_year_balt
 
ggplot(total_emissions_per_year_balt, aes(x=Year,y=Total)) + labs(x="Year", y="Total PM2.5 emission from all sources", title="Total PM2.5 emission from all sources for each year") + geom_bar(width=0.8, stat="identity", fill="#5b23ad")
```

**Total emissions in Baltimore decreased from 1999 to 2008**

3. Of the four types of sources indicated by the `type` (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for **Baltimore City**? Which have seen increases in emissions from 1999–2008?

```{r}
NEI_Balt <- NEI[NEI$fips=="24510",]
ggplot(NEI_Balt, aes(factor(year), Emissions, fill=type)) + geom_bar(stat="identity") + facet_grid(.~type) + labs(x="Year", y="Total Emission", title="Emissions in Baltimore City")
```

**In general, there is a reduction in total emissions for three types of sources (nonroad, nonpoint, onroad). For the fourth type of point, there is a significant increase in emissions in 2005 and a significant decrease in 2008, but the level of emissions is still higher than in 1999.**

4. Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

```{r}
combustion <- grepl("comb",  SCC$SCC.Level.One, ignore.case=TRUE)
coal <- grepl("coal", SCC$SCC.Level.Four, ignore.case=TRUE) 
SCC_coal_combustion <- SCC[(combustion & coal),]$SCC
NEI_coal_combustion <- NEI[NEI$SCC %in% SCC_coal_combustion,]
head(NEI_coal_combustion)
ggplot(NEI_coal_combustion, aes(factor(year), Emissions)) + geom_bar(stat="identity", fill="#5b23ad") + labs(x="Year", y="Total emission", title="Coal combustion-related sources emission")
```

**There is a reduction in emissions**

5. How have emissions from motor vehicle sources changed from 1999–2008 in **Baltimore City** (EI.Sector starts from "Mobile")?

```{r}
vehicle <- grepl("vehicle",  SCC$SCC.Level.Two, ignore.case=TRUE)
SCC_vehicle <- SCC[vehicle,]$SCC
NEI_vehicle <- NEI[NEI$SCC %in% SCC_vehicle,]
NEI_vehicle_balt <- NEI_vehicle[NEI_vehicle$fips == 24510,]
head(NEI_vehicle_balt)
ggplot(NEI_vehicle_balt, aes(factor(year), Emissions)) + geom_bar(stat="identity", fill="#5b23ad") + labs(x="Year", y="Total emission", title="Motor vehicle sources emission in Baltimore")
```

**There is a reduction in emissions**

6. Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in **Los Angeles County**, California (`fips == "06037"`). Which city has seen greater changes over time in motor vehicle emissions?

```{r}
NEI_vehicle_la <- NEI_vehicle[NEI_vehicle$fips == "06037",]
NEI_vehicle_balt_la <- rbind(NEI_vehicle_balt, NEI_vehicle_la)
cities_names <- c(`06037` = "LA", `24510` = "Baltimore")
ggplot(NEI_vehicle_balt_la, aes(factor(year), Emissions, fill=fips)) + geom_bar(stat="identity") + guides(fill="none") + facet_grid(.~fips, labeller=as_labeller(cities_names)) + labs(x="Year", y="Total Emission", title="Emissions in LA and Baltimore City")
```

**Although Baltimore has significantly lower emissions, Los Angeles has seen the largest changes in emissions (both down and down)**